CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes as ground truth for deep learning

### 2019-08-30 PaperCVFAD2019/egpaper_for_review.tex
add rmfi paper

% \iccvfinalcopy % *** Uncomment this line for the final submission

@@

\author{Yansel González Tejeda\\
Paris Lodron University of Salzburg\\
{\tt\small yansel.gonzalez-tejeda@stud.sbg.ac.at}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Helmut Mayer\\
{\tt\small helmut@cosy.sbg.ac.at}
}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
   In this paper we present CALVIS, a method to calculate chest, waist and 
   pelvis circumference from 3D human body meshes. Our motivation is to use 
   this data as ground truth for learning with convolutional neural networks 
   (CNN). Previous work had used the large scale CAESAR dataset, determined 
   $\textit{manually}$ these anthropometrical measurements or directly acquired 
   them from a person. The problem is that acquiring these data is a cost and 
   time consuming endeavor. In contrast our method can be used on 
   3D meshes automatically. First, we slice cross sections along the Y axis of 
   the mesh. Second, we calculate the cross sections length and establish a 
   signature for the body represented by the mesh. Finally, using this 
   signature we are able to calculate the chest, waist and pelvis 
   circumference by searching for critical points. We conduct two experiments. 
   In the first experiment we synthesize 10 human body meshes. Then we apply 
   CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
   results qualitatively. We observe that the measurements can indeed be used 
   to estimate the shape of a person. The second experiment serves as a 
   proof-of-concept where we use the calculated human dimensions as a ground 
   truth to train a small CNN. The idea is to establish the plausibility of our 
   approach. After having trained the network with our data, we demonstrate 
   that learning converges, achieving a 3 percent prediction error. 
   Furthermore, we make the implementation of CALVIS publicly available to 
   advance the field.
\end{abstract}

@@

Explain method here in a very succinct way. balslfas iourgioafdnsva 
ihasoifghksv kiashdikhjioapgjgf oiasjojaspofdgjaposjg aisfhdgipoashfgihasghj 
aishgfiashgiahsfighaishg ihaspioghpaioshgpiasfgpi 
pjasoip09erujfdnaihg8qzßehgiqhfe.

@@

%------------------------------------------------------------------------
\section{Approach}

@@

\subsection{Human Body Mesh Signature}
Let us consider a human body mesh $\mathcal{M}$. Our method requires that 
$\mathcal{M}$ is standing with arms raised 
parallel to the 
ground at shoulder height at a $90^\circ$ angle. In the line 
of previous work (\cite{Dibra.2016a}), we name this pose the zero (also 
normalized (rf)) pose $\vec{\theta}^*$. Additionally, we assume that the mesh 
has LSA orientation, e.g., x, y and z axis are positively directed from 
right-to-left, inferior-to-superior and posterior-to-anterior, respectively. If 
the mesh has another orientation we can always apply transformations to 
LSA-align it.

Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. Similarly, the hip 
circumference is measured often around the widest part of hips and buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves (\cite{book.compu.topo}) along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi}$ parallel to the floor to intersect the mesh. 
Since $\mathcal{M}$ is 
triangulated, the boundary of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the boundary length by specifying a point along the y-axis $j \in 
\mathbb{R}$ as
\begin{align}
\mathcal{BL}(\mathcal{M}, \boldsymbol{\pi}, j) = \sum_{i = 
1}^{i = n}s_i
\end{align}

@@

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}.}
	\label{fig:hbm_signature}
\end{figure}

@@

\subsection{Mesh Signature Extrema}

By comparing neighboring values of the mesh signature $\mathcal{MS}$ as defined 
in equation \ref{eq:mesh_signature} we find several \textit{extrema}. In 
general, we expect these \textit{extrema} to be adequate features to 
calculate the human dimensions. More specifically, we assume that:
\begin{enumerate}
	\item The global maximum $M_g$ represents the length of the largest path 
	around the arms. From a topographic point of view $M_g$ is the highest peak 
	of $\mathcal{MS}$.   
	\item The local maximum $M_{hc}$ with largest value other than $M_g$ 
	immediately posterior to it represents the hip circumference.
	\item The local minimum $M_{wc}$ posterior to $M_g$ but prior to $M_{hc}$ 
	represents waist circumference.
	\item from a topograhpic point of view $M_g$ is a peak from  	
	The local maximum $M_{cc}$ posterior to $M_g$ but prior to $M_{wc}$ 
	represents 
	chest circumference. If this local maximum does not exist we set $M_{cc}$ 
	to the value of the path $\mathcal{BL}(j_{cc})$ with largest value $M_{cc}$ 
	posterior 
	to 
	$M_g$ and difference at least $\delta$ to it but prior to $M_{wc}$	
	represents 
	chest circumference.
\end{enumerate}

@@

We conduct two experiments. In the first experiment we synthesize 10 human body 
meshes. Then we apply CALVIS to calculate chest, waist and pelvis 
circumference. We evaluate the results qualitatively by comparing our method 
with \cite{Dibra.2016b}. We observe that the 
measurements can indeed be used to estimate the shape of a person. The second 
experiment serves as a proof-of-concept where we input the calculated human 
dimensions to an artificial neural network. The idea is to establish the 
plausibility of our approach. After having trained the network with our data, 
we proof that the network is able to conduct this task.

@@

You must include your signed IEEE copyright release form when you submit
your finished paper. We MUST have this form before your paper can be
published in the proceedings.

>>>>

 \iccvfinalcopy % *** Uncomment this line for the final submission

@@

\author{Yansel González Tejeda\\
Paris Lodron University of Salzburg\\
{\tt\small yansel.gonzalez-tejeda@stud.sbg.ac.at}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Helmut Mayer\\
Paris Lodron University of Salzburg\\
{\tt\small helmut@cosy.sbg.ac.at}
}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
   In this paper we present CALVIS, a method to calculate chest, waist and 
   pelvis circumference from 3D human body meshes. Our motivation is to use 
   this data as ground truth for learning with convolutional neural networks 
   (CNN). Previous work had used the large scale CAESAR dataset, determined 
   $\textit{manually}$ these anthropometrical measurements or directly acquired 
   them from a person. The problem is that acquiring these data is a cost and 
   time consuming endeavor. In contrast our method can be used on 
   3D meshes automatically. First, we slice cross sections along the Y axis of 
   the mesh. Second, we calculate the cross sections length and establish a 
   signature for the body represented by the mesh. Finally, using this 
   signature we are able to calculate the chest, waist and pelvis 
   circumference by searching for extrema. We conduct two experiments. 
   In the first experiment we synthesize 10 human body meshes. Then we apply 
   CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
   results qualitatively. We observe that the measurements can indeed be used 
   to estimate the shape of a person. The second experiment asses the 
   plausibility of our approach where we use the calculated human dimensions as 
   a ground truth to train a small CNN. After having trained the network with 
   our data, we demonstrate that learning converges, achieving $x$ percent 
   prediction error. 
   Furthermore, we make the implementation of CALVIS publicly available to 
   advance the field.
\end{abstract}

@@

@@

%------------------------------------------------------------------------
\section{Approach}\label{sec:approach}

@@

\subsection{Human Body Mesh Signature}\label{subsec:hbm_signature}
Let us consider a human body mesh $\mathcal{M}$. Our method requires that 
$\mathcal{M}$ is standing with arms raised 
parallel to the 
ground at shoulder height at a $90^\circ$ angle. In the line 
of previous work (\cite{Dibra.2016a}), we name this pose the zero (also 
normalized (rf)) pose $\vec{\theta}^*$. Additionally, we assume that the mesh 
has LSA orientation, e.g., x, y and z axis are positively directed from 
right-to-left, inferior-to-superior and posterior-to-anterior, respectively. If 
the mesh has another orientation we can always apply transformations to 
LSA-align it.

Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. Similarly, the pelvis 
circumference is measured often around the widest part of hips and buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves (\cite{book.compu.topo}) along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi}$ parallel to the floor to intersect the mesh. 
Since $\mathcal{M}$ is 
triangulated, the boundary of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the boundary length by specifying a point along the y-axis $j \in 
\mathbb{R}$ as
\begin{align}
\mathcal{BL}(\mathcal{M}, \boldsymbol{\pi}, j) = \sum_{i = 
1}^{i = n}s_i
\end{align}

@@

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:hbm_signature}
\end{figure}

@@

\subsection{Mesh Signature Extrema}\label{subsec:mesh_signature_extrema}

By comparing neighboring values of the mesh signature $\mathcal{MS}$ as defined 
in equation \ref{eq:mesh_signature} we find several \textit{extrema}. In 
general, we expect these \textit{extrema} to be adequate features to 
calculate the human dimensions. More specifically, we assume that:
\begin{enumerate}
	\item The global maximum $M_g$ represents the length of the largest path 
	around the arms. From a topographic point of view $M_g$ is the highest peak 
	of $\mathcal{MS}$. A horizontal line at height proportion $h$ of the peak 
	intersects 
	it at two points left $p_l$ and right $p_r$.
	\item The local maximum $M_{pc}$ with largest value other than $M_g$ 
	immediately posterior to $p_r$ represents the pelvis circumference.
	\item The local minimum $M_{wc}$ posterior to $p_r$ but prior to $M_{pc}$ 
	represents waist circumference.
	\item The local maximum $M_{cc}$ posterior to $p_r$ but prior to $M_{wc}$ 
	represents 
	chest circumference. If this local maximum does not exist we set $M_{cc}$ 
	to the value of the path $\mathcal{BL}(j_{cc})$ with largest value in the 
	interval $(p_l; M_{wc})$.
\end{enumerate}

@@

We conduct two experiments. In the first experiment we synthesize eight (four 
female and four male) human body meshes using shape parameters provided by 
SURREAL \cite{varol17_surreal}.
The meshes reflect human figures characteristics such as bulky, slim, small and 
tall subjects. Then we apply CALVIS to calculate chest, waist and pelvis 
circumference. Since we do not have ground truth, we evaluate the results 
qualitatively by comparing our method 
with \cite{Dibra.2016b}.
The second 
experiment serves to asses the plausibility of our approach to use the 
synthetic data for deep learning. We input the calculated human 
dimensions to an artificial neural network. After having trained the network 
with our data, we show that learning converges.

@@

The method can be optimized. Further research must be conducted.

### 2019-08-30 PaperRMFI2019/samplepaper.tex
add rmfi paper

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{iccv}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

 \iccvfinalcopy % *** Uncomment this line for the final submission

\def\iccvPaperID{****} % *** Enter the ICCV Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ificcvfinal\pagestyle{empty}\fi

\begin{document}

%%%%%%%%% TITLE
\title{CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes 
as ground truth for deep learning}

\author{Yansel González Tejeda\\
Paris Lodron University of Salzburg\\
{\tt\small yansel.gonzalez-tejeda@stud.sbg.ac.at}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Helmut Mayer\\
Paris Lodron University of Salzburg\\
{\tt\small helmut@cosy.sbg.ac.at}
}

\maketitle
% Remove page # from the first page of camera-ready.
\ificcvfinal\thispagestyle{empty}\fi

%%%%%%%%% ABSTRACT
\begin{abstract}
   In this paper we present CALVIS, a method to calculate chest, waist and 
   pelvis circumference from 3D human body meshes. Our motivation is to use 
   this data as ground truth for learning with convolutional neural networks 
   (CNN). Previous work had used the large scale CAESAR dataset, determined 
   $\textit{manually}$ these anthropometrical measurements or directly acquired 
   them from a person. The problem is that acquiring these data is a cost and 
   time consuming endeavor. In contrast our method can be used on 
   3D meshes automatically. First, we slice cross sections along the Y axis of 
   the mesh. Second, we calculate the cross sections length and establish a 
   signature for the body represented by the mesh. Finally, using this 
   signature we are able to calculate the chest, waist and pelvis 
   circumference by searching for extrema. We conduct two experiments. 
   In the first experiment we synthesize 10 human body meshes. Then we apply 
   CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
   results qualitatively. We observe that the measurements can indeed be used 
   to estimate the shape of a person. The second experiment asses the 
   plausibility of our approach where we use the calculated human dimensions as 
   a ground truth to train a small CNN. After having trained the network with 
   our data, we demonstrate that learning converges, achieving $x$ percent 
   prediction error. 
   Furthermore, we make the implementation of CALVIS publicly available to 
   advance the field.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

\textbf{Motivation} 
Predicting 3D human body measurements from images is crucial in several 
scenarios 
like virtual try-on, animating, ergonomics, computational forensics and 
even health and mortality estimation. Researches had 
named this 
measurements body intuitive controls (\cite{Allen.2003}), biometric 
measurements (\cite{Sigal.2008}), body dimensions 
(\cite{DBLP:conf/bmvc/ChenRC11}), semantical parameters (\cite{Yang.2014}), 
traditional anthropometric 
measurements 
(\cite{Wuhrer2011}) or only ``shape" as in human shape estimation 
(\cite{Guan.2013}, \cite{Bogo:ECCV:2016}, \cite{Loper.2015}, 
\cite{Dibra.2016a}, \cite{Pishchulin.2017}). In contrast, we assume a more 
anthropometric approach 
motivated by comprehensive compendiums like Panero and Zelnik, 1979 
\cite{panero1979human}.
Throughout this paper the term human body 
dimensions 
will be used to refer to the above measurements.

@@

\subsection{Human Body Mesh Signature}\label{subsec:hbm_signature}
Let us consider a human body mesh $\mathcal{M}$. Our method requires that 
$\mathcal{M}$ is standing with arms raised 
parallel to the 
ground at shoulder height at a $90^\circ$ angle. In the line 
of previous work (\cite{Dibra.2016a}), we name this pose the zero (also 
normalized (rf)) pose $\vec{\theta}^*$. Additionally, we assume that the mesh 
has LSA orientation, e.g., x, y and z axis are positively directed from 
right-to-left, inferior-to-superior and posterior-to-anterior, respectively. If 
the mesh has another orientation we can always apply transformations to 
LSA-align it.

Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. Similarly, the pelvis 
circumference is measured often around the widest part of hips and buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves (\cite{book.compu.topo}) along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi}$ parallel to the floor to intersect the mesh. 
Since $\mathcal{M}$ is 
triangulated, the boundary of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the boundary length by specifying a point along the y-axis $j \in 
\mathbb{R}$ as
\begin{align}
\mathcal{BL}(\mathcal{M}, \boldsymbol{\pi}, j) = \sum_{i = 
1}^{i = n}s_i
\end{align}

@@

The method can be optimized. Further research must be conducted.

{\small
\bibliographystyle{ieee}
\bibliography{egbib}
}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
\usepackage{amsmath}
\usepackage{amssymb}
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, wAist and peLVIS circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yansel Gonzalez Tejeda\orcidID{0000-0003-1002-3815} \and
Helmut Mayer\orcidID{1111-2222-3333-4444}}
%
\authorrunning{Gonzalez and Mayer}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Paris Lodron University of 
	Salzburg, Kapitelgasse 4-65020 Salzburg, Austria
\url{https://www.uni-salzburg.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate chest, waist and 
pelvis circumference from 3D human body meshes. Our motivation is to use 
this data as ground truth for learning with convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset, determined 
$\textit{manually}$ these anthropometrical measurements or directly acquired 
them from a person. The problem is that acquiring these data is a cost and 
time consuming endeavor. In contrast our method can be used on 
3D meshes automatically. First, we slice cross sections along the Y axis of 
the mesh. Second, we calculate the cross sections length and establish a 
signature for the body represented by the mesh. Finally, using this 
signature we are able to calculate the chest, waist and pelvis 
circumference by searching for extrema. We conduct two experiments. 
In the first experiment we synthesize 10 human body meshes. Then we apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively. We observe that the measurements can indeed be used 
to estimate the shape of a person. The second experiment asses the 
plausibility of our approach where we use the calculated human dimensions as 
a ground truth to train a small CNN. After having trained the network with 
our data, we demonstrate that learning converges, achieving $x$ percent 
prediction error. 
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\textbf{Motivation} 
Predicting 3D human body measurements from images is crucial in several 
scenarios 
like virtual try-on, animating, ergonomics, computational forensics and 
even health and mortality estimation. Researches had 
named this 
measurements body intuitive controls (\cite{Allen.2003}), biometric 
measurements (\cite{Sigal.2008}), body dimensions 
(\cite{DBLP:conf/bmvc/ChenRC11}), semantical parameters (\cite{Yang.2014}), 
traditional anthropometric 
measurements 
(\cite{Wuhrer2011}) or only ``shape" as in human shape estimation 
(\cite{Guan.2013}, \cite{Bogo:ECCV:2016}, \cite{Loper.2015}, 
\cite{Dibra.2016a}, \cite{Pishchulin.2017}). In contrast, we assume a more 
anthropometric approach 
motivated by comprehensive compendiums like Panero and Zelnik, 1979 
\cite{panero1979human}.
Throughout this paper the term human body 
dimensions 
will be used to refer to the above measurements.

@@

\subsection{Human Body Mesh Signature}\label{subsec:hbm_signature}
Let us consider a human body mesh $\mathcal{M}$. Our method requires that 
$\mathcal{M}$ is standing with arms raised 
parallel to the 
ground at shoulder height at a $90^\circ$ angle. In the line 
of previous work \cite{Dibra.2016a}, we name this pose the zero (also 
normalized (rf)) pose $\vec{\theta}^*$. Additionally, we assume that the mesh 
has LSA orientation, e.g., x, y and z axis are positively directed from 
right-to-left, inferior-to-superior and posterior-to-anterior, respectively. If 
the mesh has another orientation we can always apply transformations to 
LSA-align it.

Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. Similarly, the pelvis 
circumference is measured often around the widest part of hips and buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves \cite{book.compu.topo} along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi}$ parallel to the floor to intersect the mesh. 
Since $\mathcal{M}$ is 
triangulated, the boundary of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the boundary length by specifying a point along the y-axis $j \in 
\mathbb{R}$ as
\begin{equation}
\mathcal{BL}(\mathcal{M}, \boldsymbol{\pi}, j) = \sum_{i = 
	1}^{i = n}s_i
\end{equation}

@@

The method can be optimized. Further research must be conducted.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{egbib}
%
